---
title: "Hands-on_Ex09"
author: "HTC"
editor: visual
date-modified: "last-modified"
format:
  html:
    toc-depth: 4
    toc-expand: true
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## **17.1 Overview**

### \*\* 17.1.1 Learning Outcome\*\*

By the end of this hands-on exercise, you will be able:

-   to import GIS polygon data into R and save them as simple feature data frame by using appropriate functions of sf package of R;

-   to import aspatial data into R and save them as simple feature data frame by using appropriate functions of sf package of R;

-   to computer accessibility measure by using Hansen’s potential model and Spatial Accessibility Measure (SAM); and

-   to visualise the accessibility measures by using tmap and ggplot2 packages.

## **17.2 The data**

-   `MP14_SUBZONE_NO_SEA_PL`: URA Master Plan 2014 subzone boundary GIS data. This data set is downloaded from data.gov.sg.

-   `hexagons`: A 250m radius hexagons GIS data. This data set was created by using [*st_make_grid()*](https://r-spatial.github.io/sf/reference/st_make_grid.html) of sf package. It is in ESRI shapefile format.

-   `ELDERCARE`: GIS data showing location of eldercare service. [This data](https://data.gov.sg/dataset/eldercare-services) is downloaded from data.gov.sg. There are two versions. One in ESRI shapefile format. The other one in Google kml file format. For the purpose of this hands-on exercise, ESRI shapefile format is provided.

-   `OD_Matrix`: a distance matrix in csv format. There are six fields in the data file. They are:

    -   `origin_id`: the unique id values of the origin (i.e. `fid` of hexagon data set.),

    -   `destination_id`: the unique id values of the destination (i.e. `fid` of `ELDERCARE` data set.),

    -   `entry_cost`: the perpendicular distance between the origins and the nearest road),

    -   `network_cost`: the actual network distance from the origin and destination,

    -   `exit_cost`: the perpendicular distance between the destination and the nearest road), and

    -   `total_cost`: the summation of `entry_cost`, `network_cost` and `exit_cost`.

All the values of the cost related fields are in **metres**.\

## **17.3 Getting Started**

```{r}
pacman::p_load(tmap, SpatialAcc, sf, 
               ggstatsplot, reshape2,
               tidyverse)
```

## **17.4 Geospatial Data Wrangling**

### **17.4.1 Importing geospatial data**

Three geospatial data will be imported from the *data/geospatial* sub-folder. They are MP14_SUBZONE_NO_SEA_PL, hexagons and ELDERCARE.

The code chunk below is used to import these three data sets shapefile by using *st_read()* of **sf** packages.

```{r}
setwd("C:/Users/HT Chen/Desktop/ISSS626/ISSS626-wp/Hands-on_Ex/Hands-on_Ex09")
mpsz <- st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_NO_SEA_PL")
```

```{r}
setwd("C:/Users/HT Chen/Desktop/ISSS626/ISSS626-wp/Hands-on_Ex/Hands-on_Ex09")
hexagons <- st_read(dsn = "data/geospatial", layer = "hexagons") 
```

```{r}
colnames(hexagons)
```

```{r}
setwd("C:/Users/HT Chen/Desktop/ISSS626/ISSS626-wp/Hands-on_Ex/Hands-on_Ex09")
eldercare <- st_read(dsn = "data/geospatial", layer = "ELDERCARE") 
```

### **17.4.2 Updating CRS information**

The code chunk below updates the newly imported *mpsz* with the correct ESPG code (i.e. 3414)

```{r}
mpsz <- st_transform(mpsz, 3414)
eldercare <- st_transform(eldercare, 3414)
hexagons <- st_transform(hexagons, 3414)
```

```{r}
st_crs(mpsz)
```

### **17.4.3 Cleaning and updating attribute fields of the geospatial data**

There are many redundant fields in the data tables of both `eldercare` and `hexagons`. The code chunks below will be used to exclude those redundant fields. At the same time, a new field called `demand` and a new field called `capacity` will be added into the data table of `hexagons` and `eldercare` sf data frame respectively. Both fields are derive using *mutate()* of **dplyr** package.

```{r}
eldercare <- eldercare %>%
  select(fid, ADDRESSPOS) %>%
  mutate(capacity = 100)
```

```{r}
hexagons <- hexagons %>%
  select(fid) %>%
  mutate(demand = 100)
```

## **17.5 Apsaital Data Handling and Wrangling**

### **17.5.1 Importing Distance Matrix**

The code chunk below uses *read_cvs()* of **readr** package to import `OD_Matrix.csv` into RStudio. The imported object is a tibble data.frame called `ODMatrix`.

```{r}
setwd("C:/Users/HT Chen/Desktop/ISSS626/ISSS626-wp/Hands-on_Ex/Hands-on_Ex09")
ODMatrix <- read_csv("data/aspatial/OD_Matrix.csv", skip = 0)
```

### **17.5.2 Tidying distance matrix**

The imported ODMatrix organised the distance matrix columnwise.

The rows represent **origins** (i.e. also know as **from** field) and the columns represent **destination** (i.e. also known as **to** field.)

The code chunk below uses *spread()* of **tidyr** package is used to transform the O-D matrix from a thin format into a fat format.

```{r}
distmat <- ODMatrix %>%
  select(origin_id, destination_id, total_cost) %>%
  spread(destination_id, total_cost)%>%
  select(c(-c('origin_id')))
```

Note: Since tidyr version 1.0 a new function called [*pivot_wider()*](https://tidyr.tidyverse.org/reference/pivot_wider.html) is introduce. You should use *pivot_wider()* instead of *spread()*

Currently, the distance is measured in metre because SVY21 projected coordinate system is used. The code chunk below will be used to convert the unit f measurement from metre to kilometre.

```{r}
distmat_km <- as.matrix(distmat/1000)
```

## **17.6 Modelling and Visualising Accessibility using Hansen Method**

### **17.6.1 Computing Hansen’s accessibility**

Now, we ready to compute Hansen’s accessibility by using *ac()* of [**SpatialAcc**](https://cran.r-project.org/web/packages/SpatialAcc/index.html) package. Before getting started, you are encourage to read the [arguments](https://cran.r-project.org/web/packages/SpatialAcc/SpatialAcc.pdf) of the function at least once in order to ensure that the required inputs are available.

The code chunk below calculates Hansen’s accessibility using *ac()* of **SpatialAcc** and *data.frame()* is used to save the output in a data frame called `acc_Handsen`.

```{r}
#acc_Hansen_0.5 <- data.frame(ac(hexagons$demand,
#                            eldercare$capacity,
#                            distmat_km, 
#                            #d0 = 50,
#                            power = 0.5, 
#                            family = "Hansen"))
```

```{r}
#acc_Hansen_1 <- data.frame(ac(hexagons$demand,
#                            eldercare$capacity,
#                            distmat_km, 
#                            d0 = 50,
#                            power = 1, 
#                            family = "Hansen"))
```

```{r}
#acc_Hansen_1.5 <- data.frame(ac(hexagons$demand,
#                            eldercare$capacity,
#                            distmat_km, 
#                            d0 = 50,
#                            power = 1.5, 
#                            family = "Hansen"))
```

```{r}
#acc_Hansen_2 <- data.frame(ac(hexagons$demand,
 #                           eldercare$capacity,
#                            distmat_km, 
#                            d0 = 50,
#                            power = 2, 
#                            family = "Hansen"))
```

```{r}
#acc_Hansen2.5 <- data.frame(ac(hexagons$demand,
#                            eldercare$capacity,
#                            distmat_km, 
#                            d0 = 50,
#                            power = 2.5, 
#                            family = "Hansen"))
```

```{r}
#tm_shape(hexagon_SAM,
#         bbox = mapex) + 
#  tm_polygons(fill = "accSAM", 
#              fill.scale = tm_scale_intervals(
#                style = "quantile",
#                n = 10,
#                values = "brewer.blues")) +
#tm_shape(eldercare) +
#  tm_dots(size = 0.3,
#          fill = "red",
#          col = "black") +
#  tm_title("Accessibility to eldercare: SAM method") +
#  tm_layout(frame = TRUE) +
#  tm_compass(type="8star", size = 2) +
#  tm_scalebar() +
#  tm_grid(alpha = 0.2)
```

```{r}
colnames(hexagons)
```

The default field name is very messy, we will rename it to `accHansen` by using the code chunk below.

```{r}

```

```{r}

```

```{r}

```

```{r}

```

### 

```{r}

```

```{r}

```

### 

```{r}

```

```{r}

```

## 

```{r}

```

### 

```{r}

```

### 

```{r}

```

```{r}

```

## 

```{r}

```

### 

```{r}

```

### 

```{r}

```

```{}
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```
