---
title: "Take-Home_Exercise_1b"
author: "HTC"
date: "September 15, 2025"
date-modified: "last-modified"
format:
  html:
    toc-depth: 4
    toc-expand: true
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Overview

### This is Take-Home Exercise 1, 1.1

# Importing Packages

```{r}
pacman::p_load(sf, tidyverse, dplyr, lubridate, httr, tmap,purrr,raster, sparr,rvest,spatstat.geom, spatstat.core)
```

# Data Wrangling

```{r}
setwd("C:/Users/HT Chen/Desktop/ISSS626/ISSS626-wp/Assignment1")
folder_path <- "data/ACRAInformationonCorporateEntities"
file_list <- list.files(path = folder_path,
                        pattern = "^ACRA*.*\\.csv$",
                        full.names =TRUE)
acra_data <- file_list %>%
  map_dfr(read_csv)
```

```{r}
setwd("C:/Users/HT Chen/Desktop/ISSS626/ISSS626-wp/Assignment1")
write_rds(acra_data,
          "data/acra_data.rds")
```

### SSIC Code(Food kiosks mainly for takeaway and delivery) within Specific Timeframe

```{r}
biz_56123 <- acra_data %>%
  dplyr::select(1:24) %>%
  filter(primary_ssic_code ==56123) %>%
  rename(date = registration_incorporation_date) %>%
  mutate(date = as.Date(date),
         YEAR = year(date),
         MONTH_NUM = month(date),
         MONTH_ABBR = month(date,label = TRUE, abbr = TRUE)) %>%
  mutate(
    postal_code = str_pad(postal_code,
    width = 6, side = "left", pad = "0")) %>% 
          filter((YEAR == 2024 & MONTH_NUM >= 1) |
         (YEAR == 2025 & MONTH_NUM <= 6))
```

```{r}
postcodes <- unique(biz_56123$postal_code)

url <- "https://onemap.gov.sg/api/common/elastic/search"

found <- data.frame()
not_found <- data.frame(postcode = character())

for (pc in postcodes) {
  query <- list(
    searchVal = pc,
    returnGeom = "Y",
    getAddrDetails = "y",
    pageNum = "1")
  res <- GET (url, query = query) 
  json <- content(res)
  if (json$found != 0) {
    df <- as.data.frame(json$results, stringsAsFactors = FALSE) 
    df$input_postcode <- pc 
    found <- bind_rows (found, df)
} else {
not_found <- bind_rows (not_found, data.frame(postcode = pc))
}}
```

```{r}
biz_56123 = biz_56123 %>% 
  left_join(found,
            by = c('postal_code' = 'POSTAL'))
```

### Geocode and Simple Point Features Conversion

```{r}
biz_geo <- biz_56123 %>%
  mutate(postal_code = str_pad(as.character(postal_code), 6, pad = "0")) %>%
  left_join(found %>%
              mutate(POSTAL = str_pad(as.character(POSTAL), 6, pad = "0")),
            by = c("postal_code" = "POSTAL"))
```

```{r}
biz_geo <- biz_geo %>%
  filter(!is.na(LONGITUDE.x), !is.na(LATITUDE.x))

```

```{r}
#colnames(biz_geo)
```

```{r}

#colSums(is.na(biz_geo))

```

```{r}
biz_geo <- biz_56123 %>%
  mutate(postal_code = str_pad(as.character(postal_code), 6, pad = "0")) %>%
  left_join(found_clean, by = c("postal_code" = "POSTAL")) %>%
  filter(!is.na(LONGITUDE.x), !is.na(LATITUDE.x))

```

```{r}
glimpse(biz_geo)

```

```{r}
biz_sf_wgs <- st_as_sf(biz_geo,
                       coords = c("LONGITUDE.x", "LATITUDE.x"),
                       crs = 4326) %>%
  st_transform(crs = 3414) %>%
  st_geometry(biz_geo)
```

```{r}
biz_ppp <- as.ppp(biz_sf_wgs)
```

```{r}
setwd("C:/Users/HT Chen/Desktop/ISSS626/ISSS626-wp/Hands-on_Ex/Hands-on_Ex02")
mpsz_sf <- st_read("data/geospatial/MasterPlan2019SubzoneBoundaryNoSeaKML (1).kml") %>% 
  st_zm(drop = TRUE, what = "ZM") %>% st_transform(crs = 3414)
```

```{r}
extract_kml_field <- function(html_text, field_name) {
  if (is.na(html_text) || html_text == "") return(NA_character_)
  
  page <- read_html(html_text)
  rows <- page %>% html_elements("tr")
  
  value <- rows %>%
    keep(~ html_text2(html_element(.x, "th")) == field_name) %>%
    html_element("td") %>%
    html_text2()
  
  if (length(value) == 0) NA_character_ else value
}
```

```{r}
mpsz_sf <- mpsz_sf %>%
  dplyr::mutate(
    REGION_N = map_chr(Description, extract_kml_field, "REGION_N"),
    PLN_AREA_N = map_chr(Description, extract_kml_field, "PLN_AREA_N"),
    SUBZONE_N = map_chr(Description, extract_kml_field, "SUBZONE_N"),
    SUBZONE_C = map_chr(Description, extract_kml_field, "SUBZONE_C")
  ) %>%
  dplyr::select(-Name, -Description) %>%
  relocate(geometry, .after = last_col())
```

```{r}
mpsz_cl <- mpsz_sf %>%
  filter(SUBZONE_N != "SOUTHERN GROUP",
         PLN_AREA_N != "WESTERN ISLANDS",
         PLN_AREA_N != "NORTH-EASTERN ISLANDS") %>%
  st_geometry(mpsz_sf)
```

```{r}
mpsz_cl <- as.owin(mpsz_cl)
```

```{r}
plot(mpsz_cl)
plot(biz_ppp, add = TRUE, pch = 16, cex = 0.4)
```

# Compute spatial and spatio-temporal KDE of the selected business types

```{r}
biz_ppp_clip <- biz_ppp[mpsz_cl]
summary(biz_ppp_clip)
```

```{r}
dens <- density(biz_ppp_clip, sigma = 500)  
plot(dens); points(biz_ppp_clip, pch = 16, cex = 0.2)
```

```{r}

t_raw <- as.Date(biz_56123$date)
t0 <- min(t_raw, na.rm = TRUE)
t_num <- as.numeric(t_raw - t0)

```

```{r}
set.seed(1234)
boot_res <- BOOT.spattemp(
  pp   = biz_ppp_clip,   # spatial ppp
  tt   = t_num,          # time vector
  tlim = range(t_num)    # time window
)

summary(boot_res)

```

```{r}
st_kde <- spattemp.density(
  pp     = biz_ppp_clip,                         
  tt     = t_num,                                
  tlim   = range(t_num, na.rm = TRUE),           
  h      = 2050,                                  
  lambda = 165                                   
)

summary(st_kde)
plot(st_kde)
```

# **Second Order Spatial Point Patterns Analysis Methods**

```{r}
Window(biz_ppp) <- mpsz_cl
```

```{r}
mean_nn <- mean(nndist(biz_ppp))


exp_nn <- 0.5 / sqrt(intensity(biz_ppp))


NNI <- mean_nn / exp_nn
NNI
```

```{r}
G <- Gest(biz_ppp)
plot(G, main="G-function (Nearest Neighbour Distribution)")
```

```{r}
G_tm.csr <- envelope(biz_ppp, Gest, nsim = 99)
```

```{r}
plot(G_tm.csr)
```

```{r}
F <- Fest(biz_ppp)
plot(F, main="F-function (Empty Space Distribution)")

```

```{r}
F_CK.csr <- envelope(biz_ppp, Fest, nsim = 99)
```

```{r}
plot(F_CK.csr)
```

```{r}
K <- Kest(biz_ppp, correction = "border")
plot(K, main="Ripley's K-function")
```

```{r}

```

```{r}
L <- Lest(biz_ppp, correction = "border")
plot(L, main="Ripley's L-function")
```

```{r}
par(mfrow=c(2,2))
plot(G, main="G-function")
plot(F, main="F-function")
plot(K, main="K-function")
plot(L, main="L-function")
par(mfrow=c(1,1))
```

# **Describe and discuss the analysis results obtained from Step 5 and 6**

## **Anaylsis:**

**First-order (KDE / hotspot evolution):**\
At *t = 0*, takeaway kiosks appear relatively dispersed, with emerging hotspots in the central and eastern regions. By *t = 353*, intensities sharpen in the eastern corridor (Bedok, Tampines, Pasir Ris) and in southern-central Singapore (Geylang, Kallang, Queenstown). The shift suggests that kiosk growth is increasingly concentrated near dense residential estates and transport interchanges, consistent with high demand for delivery services.

**Second-order (interaction):**

G-function rises above CSR expectations at very short distances, showing that kiosks are clustered at fine scales, forming dense micro-hubs around key neighborhoods.

F-function indicates slightly more clustering than CSR at intermediate distances, though gaps still exist between clusters, likely reflecting zoning and residential distribution.

Ripley’s K and L functions consistently exceed Poisson expectations at multiple scales, confirming multi-scale clustering (small local clusters embedded in larger regional structures).\
Together, the results reveal clear clustering of takeaway kiosks aligned with residential demand and delivery logistics, rather than random or evenly spread locations.

# **For each geovisualisation prepared, write a short report of not more than 150 words describing the spatio-temporal patterns revealed by the geovisualisation and statistical graphics**

## Short Report:

**(a) Spatio-temporal KDE (hotspot maps)**\
The KDE maps show that food kiosks are not randomly distributed but concentrate around dense residential and transport corridors. Over time, hotspots strengthen in eastern and central Singapore, with new growth observed in Bedok–Tampines–Pasir Ris and sustained density in southern central areas (Geylang, Kallang, Queenstown). This suggests kiosks follow population density and delivery demand hubs, while remaining sparse in the north and west.

**(b) G-function with CSR**\
The observed G(r) rises far more steeply than CSR, showing kiosks tend to have closer nearest neighbors than expected. This indicates strong micro-scale clustering, with kiosks forming competitive clusters in high-demand locations.

**(c) F-function with CSR**\
The observed F(r) departs from CSR, showing shorter-than-expected empty spaces between kiosks. This highlights that kiosks are grouped rather than evenly spread, with relatively few “service deserts” in the urban core.

**(d) Ripley’s K and L**\
Both K and L functions exceed the Poisson baseline across wide distances, confirming clustering at multiple scales. Kiosks are not just grouped locally but form regional corridors of density, shaped by housing and transport infrastructure.

# **Planning & management implications (≤500 words)**

**Supporting consumer convenience and accessibility:**\
Takeaway kiosks cluster near dense housing estates and MRT/bus interchanges, which matches demand for fast delivery and walk-in customers. Urban planners can leverage this pattern by zoning more flexible ground-floor units near transport nodes and neighborhood centers to accommodate such businesses.

**Mitigating over-concentration:**\
Strong clustering in the eastern corridor and central city suggests potential risks of over-saturation, leading to competition-driven failures, increased waste generation, and delivery congestion. Local councils should monitor kiosk density using spatial dashboards and consider introducing caps or staged licensing in already saturated areas.

**Expanding to underserved regions:**\
The north and west remain relatively sparse in kiosk density. This may reflect both lower demand and fewer available retail-ready spaces. Incentives such as subsidized rental schemes, shared kitchen facilities, or delivery-only ghost kitchens can encourage kiosk growth in these underserved areas, ensuring more equitable food access and reducing delivery time inequalities.

**Infrastructure and logistics:**\
Dense kiosk clusters create logistical challenges for delivery drivers (traffic congestion, parking stress, and safety issues). Planners can respond by providing designated delivery bays, micro-logistics hubs, and shared cold-storage facilities near MRT stations or malls in high-density corridors.

**Resilience and public health:**\
Clusters of small F&B outlets are more vulnerable to supply chain disruptions and health risks (e.g., hygiene lapses or foodborne illness outbreaks). Authorities can reduce systemic risks by providing cluster-wide safety inspections, shared waste management services, and hygiene training programs.

**Data-driven monitoring:**\
First-order and second-order analyses can be repeated quarterly to track whether kiosk clustering is deepening or dispersing. Linking business registry data with delivery platform demand patterns would further refine insights, enabling predictive planning (e.g., anticipating emerging hotspots).

**Bottom line:**\
Food kiosks for takeaway and delivery in Singapore show strong spatial clustering aligned with housing and transit demand. Policy should focus on supporting these high-demand clusters with logistics infrastructure and hygiene safeguards, while also encouraging dispersal to underserved areas. This dual approach will ensure both efficiency and resilience in Singapore’s urban food service ecosystem.
